RDF Validation Requirements - Evaluation and Logical Underpinning

### 2015-01-05 paper/RDF validation requirements/RDF validation requirements.tex
finalize RDF validation requirements paper

\ms{Minimum Unqualified Cardinality Restrictions on Properties}
corresponds to the requirements
\ms{R-81-MINIMUM-UNQUALIFIED-CARDINALITY-ON-PROPERTIES} and
\ms{R-211-CARDINALITY-CONSTRAINTS}.
$\leq n R. \top$ is the minimum unqualified cardinality restriction where $n \in \mathbb{N}$ (written $\leq  n R$ in short).
A minimum cardinality restrictions contains all those individuals that are connected by a property to at least n different individuals/literals 
that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. 
If the data range is missing, it is taken to be rdfs:Literal.

@@

\begin{itemize}
	\item R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES
  \item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item A minimum cardinality expression ObjectMinCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at least n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A minimum cardinality expression DataMinCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at least n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated.
\end{itemize}
Definition:\\

$\leq n R. C$ is minimum qualified cardinality restriction where $n \in \mathbb{N}$.

@@

\begin{itemize}
	\item set of individuals that are children of at least two parents
\end{itemize}

\begin{ex}
# OWL 2
# -----
owl:Thing
    a owl:Restriction ;
    owl:minQualifiedCardinality "2"^^xsd:nonNegativeInteger ;
    owl:onProperty childOf ;
    owl:onClass Parent .
\end{ex}

@@

\begin{ex}
# ShEx
# ----
:Jedi {
    :attitude ('good') }
:JediStudent {
    & :Jedi ,
    :studentOf @:Jedi{1} }
:JediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{1,2} }
:SuperJediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{3,} }
\end{ex}

@@

\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:SuperJediMaster’ data shape
# ------------------------------------------------------
:Yoda

# Individuals matching the ’:JediMaster’ data shape
# -------------------------------------------------
:Obi-Wan
\end{ex}

@@

\begin{itemize}
	\item R-82-MAXIMUM-UNQUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item A maximum cardinality expression ObjectMaxCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at most n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A maximum cardinality expression DataMaxCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at most n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item unqualified: CE respective DR is NOT stated. 
\end{itemize}

Definition:\\

$\geq n R. \top$ is maximum unqualified cardinality restriction where $n \in \mathbb{N}$ (written $\geq  n R$ in short).
\subsection{Simple Example}

\begin{ex}
# ShEx
# ----
:Jedi {
    :attitude ('good') }
:JediStudent {
    & :Jedi ,
    :studentOf {}{1} }
:JediMaster {
    & :Jedi ,
    :mentorOf {}{1,2} }
:SuperJediMaster {
    & :Jedi ,
    :mentorOf {}{3,} }
\end{ex}

\begin{itemize}
	\item Jedis have the attitude 'good'
	\item Jedi students are students of exactly 1 resource
	\item Jedi masters are mentoring at least 1 and at most 2 resources
	\item Super Jedi masters are mentoring at least 3 resources  
\end{itemize}
\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:JediMaster’ data shape
# -------------------------------------------------
:Obi-Wan
\end{ex}

@@

\begin{itemize}
	\item R-76-MAXIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item A maximum cardinality expression ObjectMaxCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to at most n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing.
	\item A maximum cardinality expression DataMaxCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to at most n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated. 
\end{itemize}
Definition:\\

$\leq n R. C$ is maximum qualified cardinality restriction where $n \in \mathbb{N}$.

@@

%\tb{ToDO: check DL}

\begin{itemize}
	\item set of individuals that are children of at most two parents
\end{itemize}

@@

\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:JediMaster’ data shape
# -------------------------------------------------
:Obi-Wan
\end{ex}

%\tb{ToDO: add DL}

@@

\begin{itemize}
	\item R-80-EXACT-UNQUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item An exact cardinality expression ObjectExactCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to exactly n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing. Such an expression is actually equivalent to the expression ObjectIntersectionOf( ObjectMinCardinality( n OPE CE ) ObjectMaxCardinality( n OPE CE ) ).
	\item An exact cardinality expression DataExactCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to exactly n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item unqualified: CE respective DR is NOT stated. 
\end{itemize}

Definition:\\

$\geq n R. \top \sqcap \leq n R. \top $ is exact unqualified cardinality restriction where $n \in \mathbb{N}$. ($\top$ can be dropped in short writing.)
\subsection{Simple Example}

@@

\begin{ex}
# ShEx
# ----
:JediStudent {
    :studentOf {}{1} }
\end{ex}

\begin{ex}
# ReSh
# ----
:JediStudent a rs:ResourceShape ;
    rs:property [
        rs:name "studentOf" ;
        rs:propertyDefinition :studentOf ;
        rs:valueShape [ a rs:ResourceShape] ;
        rs:occurs rs:Exactly-one ; ] .
\end{ex}

@@

\begin{ex}
# data
# ----
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:JediStudent’ data shape
# --------------------------------------------------
:MaceWindu :Obi-Wan :Anakin :Luke
\end{ex}

@@

\begin{itemize}
	\item R-74-EXACT-QUALIFIED-CARDINALITY-ON-PROPERTIES
	\item R-211-CARDINALITY-CONSTRAINTS
\end{itemize}

description:

\begin{itemize}
	\item An exact cardinality expression ObjectExactCardinality( n OPE CE ) consists of a nonnegative integer n, an object property expression OPE, and a class expression CE, and it contains all those individuals that are connected by OPE to exactly n different individuals that are instances of CE. If CE is missing, it is taken to be owl:Thing. Such an expression is actually equivalent to the expression ObjectIntersectionOf( ObjectMinCardinality( n OPE CE ) ObjectMaxCardinality( n OPE CE ) ).
	\item An exact cardinality expression DataExactCardinality( n DPE DR ) consists of a nonnegative integer n, a data property expression DPE, and a unary data range DR, and it contains all those individuals that are connected by DPE to exactly n different literals in DR. If DR is not present, it is taken to be rdfs:Literal. 
	\item qualified: CE respective DR is stated. 
\end{itemize}

Definition:\\

$\geq n R. C \sqcap \leq n R. C $ is exact unqualified cardinality restriction where $n \in \mathbb{N}$.
\subsection{Simple Example}

@@

\begin{ex}
# ShEx
# ----
:Jedi {
    :attitude ('good') }
:JediStudent {
    :studentOf @:Jedi{1} }
\end{ex}

\begin{ex}
# ReSh
# ----
:Jedi a rs:ResourceShape ;
    rs:property [
        rs:name "attitude" ;
        rs:propertyDefinition :attitude ;
        rs:allowedValue "good" ;
        rs:occurs rs:Exactly-one ;
    ] .
:JediStudent a rs:ResourceShape ;
    rs:property [
        rs:name "studentOf" ;
        rs:propertyDefinition :studentOf ;
        rs:valueShape :Jedi ;
        rs:occurs rs:Exactly-one ;
    ] .
\end{ex}

@@

%\tb{ToDO: check DL}

\begin{itemize}
	\item every person is a child of exactly two parents
\end{itemize}

>>>>

\ms{Minimum Unqualified Cardinality Restrictions on Properties}
corresponds to the requirements
\ms{R-81-MINIMUM-UNQUALIFIED-CARDINALITY-ON-PROPERTIES} and
\ms{R-211-CARDINALITY-CONSTRAINTS}.
$\leq n R. \top$ is the minimum unqualified cardinality restriction where $n \in \mathbb{N}$ (written $\leq  n R$ in short).
A minimum cardinality restrictions contains all those individuals that are connected by a property to at least n different individuals/literals 
that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. 
If the data range is missing, it is taken to be rdfs:Literal.
For unqualified cardinality restrictions, classes respective data ranges are not stated.

@@

\ms{Minimum Qualified Cardinality Restrictions on Properties} corresponds to the requirements
\ms{R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES} and \ms{R-211-CARDINALITY-CONSTRAINTS}
A minimum cardinality restrictions contains all those individuals that are connected by a property to at least n different individuals/literals 
that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. 
If the data range is missing, it is taken to be rdfs:Literal.
For qualified cardinality restrictions, classes respective data ranges are stated.
$\leq n R. C$ is a minimum qualified cardinality restriction where $n \in \mathbb{N}$.

@@

Set of individuals that are children of at least two parents.

\begin{ex}
# OWL 2:
owl:Thing
    a owl:Restriction ;
    owl:minQualifiedCardinality "2"^^xsd:nonNegativeInteger ;
    owl:onProperty childOf ;
    owl:onClass Parent .
\end{ex}

@@

\begin{ex}
# ShEx:
:Jedi {
    :attitude ('good') }
:JediStudent {
    & :Jedi ,
    :studentOf @:Jedi{1} }
:JediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{1,2} }
:SuperJediMaster {
    & :Jedi ,
    :mentorOf @:Jedi{3,} }
\end{ex}

@@

\begin{ex}
# Individuals matching the ’:SuperJediMaster’ data shape:
:Yoda

# Individuals matching the ’:JediMaster’ data shape:
:Obi-Wan
\end{ex}

@@

\ms{Maximum Unqualified Cardinality Restrictions on Properties} corresponds to the requirements 
\ms{R-82-MAXIMUM-UNQUALIFIED-CARDINALITY-ON-PROPERTIES} and \ms{R-211-CARDINALITY-CONSTRAINTS}.
A maximum cardinality restriction contains all those individuals that are connected by a property to at most n different individuals/literals that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. If the data range is not present, it is taken to be rdfs:Literal.
Unqualified means that the class respective the data range is not stated. 
$\geq n R. \top$ is a maximum unqualified cardinality restriction where $n \in \mathbb{N}$ (written $\geq  n R$ in short).
\subsection{Simple Example}

\begin{itemize}
	\item Jedis have the attitude 'good'
	\item Jedi students are students of exactly 1 resource
	\item Jedi masters are mentoring at least 1 and at most 2 resources
	\item Super Jedi masters are mentoring at least 3 resources  
\end{itemize}
\begin{ex}
# data:
:Yoda 
    :attitude 'good' ;
    :mentorOf :MaceWindu , :Obi-Wan , :Luke .
:MaceWindu
    :attitude 'good' ;
    :studentOf :Yoda .
:Obi-Wan 
    :attitude 'good' ;
    :studentOf :Yoda ;
    :mentorOf :Anakin .
:Anakin
    :attitude 'good' ; 
    :studentOf :Obi-Wan .
:Luke
    :attitude 'good' ;
    :studentOf :Yoda .
\end{ex}

\begin{ex}
# Individuals matching the ’:JediMaster’ data shape:
:Obi-Wan
\end{ex}

@@

\ms{Maximum Qualified Cardinality Restrictions on Properties} corresponds to the requirements
\ms{R-76-MAXIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES} and
\ms{R-211-CARDINALITY-CONSTRAINTS}.
A maximum cardinality restriction contains all those individuals that are connected by a property to at most n different individuals/literals that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. If the data range is not present, it is taken to be rdfs:Literal.
Qualified means that the class respective the data range is stated. 
$\leq n R. C$ is a maximum qualified cardinality restriction where $n \in \mathbb{N}$.

@@

Set of individuals that are children of at most two parents.

@@

\begin{ex}
# Individuals matching the ’:JediMaster’ data shape:
:Obi-Wan
\end{ex}

%\noindent $mentorOf$ and $studentOf$ are taken to be inverse properties:\\

@@

\ms{Exact Unqualified Cardinality Restrictions on Properties} corresponds to the requirements
\ms{R-80-EXACT-UNQUALIFIED-CARDINALITY-ON-PROPERTIES} and
\ms{R-211-CARDINALITY-CONSTRAINTS}.
An exact cardinality restriction contains all those individuals that are connected by a property to exactly n different individuals that are instances of a particular class or data range. 
If the class is missing, it is taken to be owl:Thing. 
If the data range is not present, it is taken to be rdfs:Literal.
Unqualified means that the class respective data range is not stated. 
$\geq n R. \top \sqcap \leq n R. \top $ is an exact unqualified cardinality restriction where $n \in \mathbb{N}$.

@@

\begin{ex}
# ShEx:
:JediStudent {
    :studentOf {}{1} }
\end{ex}

\begin{ex}
# ReSh:
:JediStudent a rs:ResourceShape ;
    rs:property [
        rs:name "studentOf" ;
        rs:propertyDefinition :studentOf ;
        rs:valueShape [ a rs:ResourceShape] ;
        rs:occurs rs:Exactly-one ; ] .
\end{ex}

@@

\begin{ex}
# Individuals matching the ’:JediStudent’ data shape:
:MaceWindu :Obi-Wan :Anakin :Luke
\end{ex}

@@

\ms{Exact Qualified Cardinality Restrictions on Properties} corresponds to the requirements
\ms{R-74-EXACT-QUALIFIED-CARDINALITY-ON-PROPERTIES} and \\
\ms{R-211-CARDINALITY-CONSTRAINTS}.
An exact cardinality restriction contains all those individuals that are connected by a property to exactly n different individuals that are instances of a particular class or data range. 
If the class is missing, it is taken to be owl:Thing. 
If the data range is not present, it is taken to be rdfs:Literal.
Qualified means that the class respective data range is stated. 
$\geq n R. C \sqcap \leq n R. C $ is an exact qualified cardinality restriction where $n \in \mathbb{N}$.

@@

\begin{ex}
# ShEx:
:Jedi {
    :attitude ('good') }
:JediStudent {
    :studentOf @:Jedi{1} }
\end{ex}

\begin{ex}
# ReSh:
:Jedi a rs:ResourceShape ;
    rs:property [
        rs:name "attitude" ;
        rs:propertyDefinition :attitude ;
        rs:allowedValue "good" ;
        rs:occurs rs:Exactly-one ;
    ] .
:JediStudent a rs:ResourceShape ;
    rs:property [
        rs:name "studentOf" ;
        rs:propertyDefinition :studentOf ;
        rs:valueShape :Jedi ;
        rs:occurs rs:Exactly-one ;
    ] .
\end{ex}

@@

Every person is a child of exactly two parents.
